### 2018-06-25 pubs/itp-2018/sections/metis-language-and-proofs.tex
Attempt to realocate the resolve rule after simplify

      \texttt{conjunct}
      &Takes a formula from a conjunction
      &\ref{thm:conjunct}
      \\

      \texttt{resolve}
      &A general form of the resolution theorem
      &\ref{thm:resolve}
      \\

      \texttt{canonicalize}
      &Normalization of the formula
      &\ref{thm:canonicalize}
      \\

      \texttt{clausify}
      &Performs clausification
      &\ref{thm:clausify}
      \\

      \texttt{simplify}
      &Simplify definitions and theorems
      &\ref{thm:simplify}
      \\[1ex]
    \hline
    \end{tabular}}
  \end{center}
\end{table}

>>>>

      \texttt{conjunct}
      &Take a formula from a conjunction
      &\ref{thm:conjunct}
      \\

      \texttt{canonicalize}
      &Normalize a formula
      &\ref{thm:canonicalize}
      \\

      \texttt{clausify}
      &Perform clausification
      &\ref{thm:clausify}
      \\

      \texttt{simplify}
      &Simplify definitions and theorems
      &\ref{thm:simplify}
      \\

      \texttt{resolve}
      &Apply one form of the resolution theorem
      &\ref{thm:resolve}
      \\[1ex]

    \hline
    \end{tabular}}
  \end{center}
\end{table}

### 2018-06-25 pubs/itp-2018/sections/proof-reconstruction.tex
Attempt to realocate the resolve rule after simplify

\subfile{sections/proof-reconstruction/conjunct.tex}
\subfile{sections/proof-reconstruction/resolve.tex}
\subfile{sections/proof-reconstruction/canonicalize.tex}
\subfile{sections/proof-reconstruction/clausify.tex}
\subfile{sections/proof-reconstruction/simplify.tex}

@@

We have finished the formalization of every inference rule in
a \Metis derivation, we are able to justify step-by-step any proof
for a problem in propositional logic.
For instance, we tested successfully the translation by \Athena
jointly with the \Agda formalizations of the rules mentioned above
with more than eighty representative theorems in propositional
logic. An interested reader can test the
problems~\cite{Prieto-Cubides2017} in the \Athena tool
repository~\cite{Athena}.

>>>>

\subfile{sections/proof-reconstruction/conjunct.tex}
\subfile{sections/proof-reconstruction/canonicalize.tex}
\subfile{sections/proof-reconstruction/clausify.tex}
\subfile{sections/proof-reconstruction/simplify.tex}
\subfile{sections/proof-reconstruction/resolve.tex}

@@

We have finished the formalization of every inference rule in a \Metis
derivation, we are able to justify step-by-step any proof for a problem in
propositional logic. For instance, we tested successfully the translation by
\Athena jointly with the \Agda formalizations of the rules mentioned above with
more than eighty representative theorems in propositional logic. An interested
reader can reproduce the testing of all problems from \cite{Prieto-Cubides2017}
with the sources in the \Athena tool repository~\cite{Athena}.

### 2018-06-25 pubs/itp-2018/sections/proof-reconstruction/canonicalize.tex
[ paper ] all proofs have been removed.
Attempt to realocate the resolve rule after simplify

@@

The \canonicalize rule also removes inside of the formula
the tautologies listed in~Fig.~\ref{fig:redundancies}.

\begin{figure}
  \[
    \begin{bprooftree}
      \AxiomC{$Γ ⟝ φ ∨ ⊥$}
      \UnaryInfC{$Γ ⟝ φ$}
    \end{bprooftree}
    \qquad
    \begin{bprooftree}
      \AxiomC{$Γ ⟝ φ ∨ ⊤$}
      \UnaryInfC{$Γ ⟝ ⊤$}
    \end{bprooftree}
    \qquad
    \begin{bprooftree}
      \AxiomC{$Γ ⟝ φ ∨ ¬ φ$}
      \UnaryInfC{$Γ ⟝ ⊤$}
    \end{bprooftree}
    \qquad
    \begin{bprooftree}
      \AxiomC{$Γ ⟝ φ ∨ φ$}
      \UnaryInfC{$Γ ⟝ φ$}
    \end{bprooftree}
  \]

  \[
    \begin{bprooftree}
      \AxiomC{$Γ ⟝ φ ∧ ⊥$}
      \UnaryInfC{$Γ ⟝ ⊥$}
    \end{bprooftree}
    \qquad
    \begin{bprooftree}
      \AxiomC{$Γ ⟝ φ ∧ ⊤$}
      \UnaryInfC{$Γ ⟝ φ$}
    \end{bprooftree}
    \qquad
    \begin{bprooftree}
      \AxiomC{$Γ ⟝ φ ∧ φ$}
      \UnaryInfC{$Γ ⟝ φ$}
    \end{bprooftree}
    \qquad
    \begin{bprooftree}
      \AxiomC{$Γ ⟝ φ ∧ ¬ φ$}
      \UnaryInfC{$Γ ⟝ ⊥$}
    \end{bprooftree}
  \]
  \caption{Theorems to remove inside of a formula by the
  \canonicalize rule.}
\label{fig:redundancies}
\end{figure}

The \canonicalize rule is used by \Metis to introduce the subgoals in
their refutation proofs but also helps to simplify formulas at
intermediate steps in the derivations. As far as we know, the
\canonicalize rule implements a conjunctive normal form conversion
with simplifications of tautologies. Otherwise, when
an axiom, definition or hypothesis is needed to prove some goal, this
rule gets the negative normal form of the formula.

To reconstruct the \canonicalize rule, we adapted some ideas from the
\Metis source code. The presentation of this reconstruction is as
follows. We firstly describe functions to remove tautologies inside
of the formula. After, we present the negative normal form conversion
in~Lemma~\ref{lem:nnf} and the conjunctive normal form in
Lemma~\ref{lem:cnf}. At the end of this section, we state
Theorem~\ref{thm:canonicalize} to reconstruct the \canonicalize rule.

\begin{notation}
In a disjunction, $φ ≡ φ₁ ∨ φ₂ ∨ \cdots ∨ φₙ$, we say $ψ ∈_{∨} φ$,
if there is some $i = 1, \cdots, n$ such that $ψ ≡ φᵢ$.
Note that $ψ ∈_{∨} φ$ is another representation for the equality
$ψ ≡ \freorder_{∨}~φ~ψ$.
\end{notation}

In right-associative disjunctions we remove the tautologies
in Fig.~\ref{fig:or-redundancies} using Lemma~\ref{lem:canon-or}.
We assume the formulas to be right-associative unless otherwise
stated.

\begin{figure}
\begin{equation*}
\begin{bprooftree}
  \AxiomC{$Γ ⟝ φ ∨ ⊥$}
  \UnaryInfC{$Γ ⟝ φ$}
\end{bprooftree}\qquad
\begin{bprooftree}
  \AxiomC{$Γ ⟝ φ ∨ ⊤$}
  \UnaryInfC{$Γ ⟝ ⊤$}
\end{bprooftree}\qquad
\begin{bprooftree}
  \AxiomC{$Γ ⟝ φ ∨ φ$}
  \UnaryInfC{$Γ ⟝ φ$}
\end{bprooftree}
\begin{bprooftree}
  \AxiomC{$Γ ⟝ φ ∨ ¬~φ$}
  \UnaryInfC{$Γ ⟝ ⊤$}
\end{bprooftree}\qquad
\end{equation*}
\caption{Theorems to remove inside of a disjunction by
the \canonicalize rule.}
\label{fig:or-redundancies}
\end{figure}

\begin{mainlemma}
  \label{lem:canon-or}

  Let $φ : \Prop$ be a right-associative formula. If $Γ~⟝~φ$
  then $Γ~⟝~\fcanon_{∨}~φ$, where

\begin{equation}
\label{eq:canon-or}
\begin{aligned}
 &\hspace{.495mm}\fcanon_{∨} : \NProp \to \NProp\\
 &\begin{array}{lll}
   \fcanon_{∨} &(⊥ ∨ φ)     &= \fcanon_{∨}~φ \\
   \fcanon_{∨} &(φ ∨ ⊥)     &= \fcanon_{∨}~φ \\
   \fcanon_{∨} &(⊤ ∨ φ)     &= ⊤\\
   \fcanon_{∨} &(φ ∨ ⊤)     &= ⊤\\
   % \fcanon_{∨} &(¬ φ₁ ∨ φ₂) &=
   %      \begin{cases}
   %       \fcanon_{∨}~φ₂,        &\text{ if } ¬ φ₁ ∈_{∨} φ₂;\\
   %       ⊤,                     &\text{ if } φ₁ ∈_{∨} φ₂;\\
   %       ⊤,                     &\text{ if } \fcanon_{∨}~φ₂≡ ⊤;\\
   %       ¬ φ₁,                  &\text{ if } \fcanon_{∨}~φ₂≡ ⊥;\\
   %       ¬ φ₁ ∨ \fcanon_{∨}~φ₂, &\text{ otherwise.}
   %      \end{cases}\\
   \fcanon_{∨} &(φ₁ ∨ φ₂)   &=
    \begin{cases}
     ⊤,                     & \text{ if } φ₁ ≡ ¬ ψ \text{ for some }ψ : \Prop \text{ and } ψ ∈_{∨} φ₂;\\
     ⊤,                     & \text{ if } (¬ φ₁) ∈_{∨} φ₂;\\
     \fcanon_{∨}~φ₂,        & \text{ if } φ₁ ∈_{∨} φ₂;\\
     ⊤,                     & \text{ if } \fcanon_{∨}~φ₂~ ≡ ⊤;\\
     φ₁,                    & \text{ if } \fcanon_{∨}~φ₂~ ≡ ⊥;\\
     φ₁ ∨ \fcanon_{∨}~φ₂,   & \text{ otherwise.}
    \end{cases}\\
   \fcanon_{∨} &φ           &= φ.
  \end{array}
\end{aligned}
\end{equation}
\end{mainlemma}

Now, we have removed tautologies in disjunctions by applying the
$\fcanon_{∨}$ function. In a similar way, we define the $\fcanon_{∧}$
function to work with conjunctions.

\begin{notation}
In a conjunction, $φ ≡ φ₁ ∧ φ₂ ∧ \cdots ∧ φₙ$, we say
$ψ ∈_{∧} φ$, if there is some $i = 1, \cdots, n$ such that $ψ ≡ φᵢ$.
Note that $ψ ∈_{∧} φ$ is another representation of
the equality $ψ ≡ \fconjunct~φ~ψ$.
\end{notation}

In right-associative conjunctions we remove the tautologies
in Fig.~\ref{fig:and-redundancies} using Lemma~\ref{lem:canon-and}.

\begin{figure}
\begin{equation*}
\begin{bprooftree}
  \AxiomC{$Γ ⟝ φ ∧ ⊤$}
  \UnaryInfC{$Γ ⟝ φ$}
\end{bprooftree}\qquad
\begin{bprooftree}
  \AxiomC{$Γ ⟝ φ ∧ ⊥$}
  \UnaryInfC{$Γ ⟝ ⊥$}
\end{bprooftree}\qquad
\begin{bprooftree}
  \AxiomC{$Γ ⟝ φ ∧ φ$}
  \UnaryInfC{$Γ ⟝ φ$}
\end{bprooftree}\qquad
\begin{bprooftree}
  \AxiomC{$Γ ⟝ φ ∧ ¬~φ$}
  \UnaryInfC{$Γ ⟝ ⊥$}
\end{bprooftree}
\end{equation*}
\caption{Conjunction tautologies.}
\label{fig:and-redundancies}
\end{figure}

\begin{mainlemma}
  \label{lem:canon-and}

  Let $φ : \Prop$ be a right-associative formula. If $Γ~⟝~φ$
  then $Γ~⟝~\fcanon_{∧}~φ$, where

  \begin{equation}
   \label{eq:canon-and}
    \begin{aligned}
     &\hspace{.495mm}\fcanon_{∧} : \NProp \to \NProp\\
      &\begin{array}{lll}
        \fcanon_{∧} &(⊥ ∧ φ)     &= ⊥  \\
        \fcanon_{∧} &(φ ∧ ⊥)     &= ⊥  \\
        \fcanon_{∧} &(⊤ ∧ φ)     &= \fcanon_{∧}~φ \\
        \fcanon_{∧} &(φ ∧ ⊤)     &= \fcanon_{∧}~φ \\
        % \fcanon_{∧} &(¬ φ₁ ∧ φ₂) &=
        %   \begin{cases}
        %     ⊥,                     &\text{ if } φ₁ ∈_{∧} φ₂;\\
        %     \fcanon_{∧}~φ₂,        &\text{ if } ¬ φ₁ ∈_{∧} φ₂;\\
        %     ⊥,                     &\text{ if } \fcanon_{∧}~φ₂~≡ ⊥;\\
        %      \fcanon_{∧}~φ₂,       &\text{ if } \fcanon_{∧}~φ₂~≡ ⊤;\\
        %     ¬ φ₁ ∧ \fcanon_{∧}~φ₂, &\text{ otherwise.}
        %   \end{cases}\\
        \fcanon_{∧} &(φ₁ ∧ φ₂) &=
          \begin{cases}
            ⊥,                   & \text{ if } φ₁ ≡ ¬ ψ \text{ for some }ψ : \Prop \text{ and } ψ ∈_{∧} φ₂;\\
            ⊥,                   & \text{ if } (¬ φ₁) ∈_{∧} φ₂;\\
            \fcanon_{∧}~φ₂,      & \text{ if } φ₁ ∈_{∧} φ₂;\\
            φ₁,                  & \text{ if } \fcanon_{∧}~φ₂~≡ ⊤;\\
            ⊥,                   & \text{ if } \fcanon_{∧}~φ₂~≡ ⊥;\\
            φ₁ ∧ \fcanon_{∧}~φ₂, &\text{ otherwise.}
          \end{cases}\\
        \fcanon_{∧} &φ         &= φ.
       \end{array}
    \end{aligned}
    \end{equation}
\end{mainlemma}

Now, we are ready to define the negative normal form of a formula with
simplifications by applying to it the $\fnnf$ function defined in
Lemma~\ref{lem:nnf} (see more details in Appendix~\ref{app:polarity-for-propositions}).
This definition is mainly based on the \Metis
source code to normalize formulas. To define such a function in type
theory we used bounded recursion as we describe in
Section~\ref{ssec:structural-recursion}.

@@

\begin{mainlemma}
  \label{lem:nnf}
  If $Γ ⊢ φ$ then $Γ ⟝ \fnnf~φ$, where
  \begin{align*}
   \begin{split}
     &\fnnf : \NProp \to \NProp\\
     &\fnnf~φ = \fnnf_{1}~φ~(\fnnf_{cm}~φ).
   \end{split}
  \end{align*}
  The $\fnnf_{cm}$  complexity measure function
  is defined in Appendix~\ref{app:polarity-for-propositions}.
\end{mainlemma}

@@

\begin{proof}
  Composition of Lemma~\ref{lem:dist} and Lemma~\ref{lem:nnf}.
\end{proof}

@@

\begin{proof}\hspace{3mm}
\begin{itemize}
\item[∙] Case $φ ≡ ψ$. By substitution theorem we conclude $Γ ⊢ ψ$.
\item[∙] Case $ψ ≡ \fnnf~φ$.
\begin{equation*}
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ φ$}
    \RightLabel{Lemma~\ref{lem:nnf}}
    \UnaryInfC{$Γ ⊢~\fnnf~φ$}
    \AxiomC{$ψ ≡ \fnnf~φ$}
    \RightLabel{\fsubst}
    \BinaryInfC{$Γ ⊢ ψ$}
  \end{bprooftree}
\end{equation*}

\item[∙] Case $\fcnf~ψ ≡ \freorder_{∧∨}~(\fcnf~φ)~(\fcnf~ψ)$.
  \begin{equation*}
    \begin{bprooftree}
      \AxiomC{$Γ ⊢ φ$}
      \RightLabel{Lemma~\ref{lem:cnf}}
      \UnaryInfC{$Γ ⊢ \fcnf~φ$}
      \RightLabel{Lemma~\ref{lem:reorder-and-or}}
      \UnaryInfC{$Γ ⊢ \freorder_{∧∨}~(\fcnf~φ)~(\fcnf~ψ)$}
      \AxiomC{$\fcnf~ψ~≡~\freorder_{∧∨}~(\fcnf~φ)~(\fcnf~ψ)$}
      \RightLabel{\fsubst}
      \BinaryInfC{$Γ ⊢ \fcnf~ψ$}
      \RightLabel{Lemma~\ref{lem:cnf-inv}}
      \UnaryInfC{$Γ ⊢ ψ$}
    \end{bprooftree}
  \end{equation*}
\end{itemize}
\end{proof}

>>>>

% ----------------------------------------------------------------------------
% -- Preliminars --
% ----------------------------------------------------------------------------

\begin{mydefinition}

A \emph{literal} is a propositional variable (positive literal) or a
negation of a propositional variable (negative literal).

\end{mydefinition}

\begin{notation}
\Lit is a synonym of \Prop to refer to literals.
\end{notation}

\begin{mydefinition}

The \emph{negative normal form} of a formula is the logical
equivalent version of it in which negations appear only in the
literals and the formula does not contain any implications.

\end{mydefinition}

\begin{mydefinition}

The \emph{conjunctive normal form} of a formula also called clausal
normal form is the logical equivalent version expressed as a
conjunction of clauses where a \emph{clause} is the disjunction of
zero or more literals.
\end{mydefinition}

\begin{remark} The proof-reconstruction we are exposing needs a mechanism to
test logic equivalence between propositions in some cases. This task is not
trivial since we left out the semantics to treat only the syntax aspects of the
propositional formulas, and therefore we have to manipulate in somehow the
formulas to see if they are logically equivalent.

We address this challenge by converting the formulas to their conjunctive
normal. We check if their normal forms are equal and if they are not, we try to
reorder the first formula to match with the second one to fulfill the
syntactical equality. In~\cite{Prieto-Cubides2017a}, we make a description of a
set of functions and lemmas for such a purpose. For instance, we define
functions of ($\Source \to  \Target \to \Prop$) type following the pattern in
Example~\ref{ex:inference-rule-pattern} like the $\freorder_{∨}$ function. This
function try to fix the order of a \emph{premise} disjunction given by following
the order about the \emph{conclusion} disjunction, the target. Similarly, we
define the function $\freorder_{\wedge}$ for conjunctions and the function
$\freorder_{\wedge\vee}$ for conjunctive normal forms.

\end{remark}

% ----------------------------------------------------------------------------
% -- CANONICALIZE --
% ----------------------------------------------------------------------------

@@

\begin{figure}
  \[
    \text{Disjunction:}
    \begin{bprooftree}
      \AxiomC{$Γ ⟝ φ ∨ ⊥$}
      \UnaryInfC{$Γ ⟝ φ$}
    \end{bprooftree}
    \qquad
    \begin{bprooftree}
      \AxiomC{$Γ ⟝ φ ∨ ⊤$}
      \UnaryInfC{$Γ ⟝ ⊤$}
    \end{bprooftree}
    \qquad
    \begin{bprooftree}
      \AxiomC{$Γ ⟝ φ ∨ ¬ φ$}
      \UnaryInfC{$Γ ⟝ ⊤$}
    \end{bprooftree}
    \qquad
    \begin{bprooftree}
      \AxiomC{$Γ ⟝ φ ∨ φ$}
      \UnaryInfC{$Γ ⟝ φ$}
    \end{bprooftree}
  \]

  \[
    \text{Conjunction:}
    \begin{bprooftree}
      \AxiomC{$Γ ⟝ φ ∧ ⊥$}
      \UnaryInfC{$Γ ⟝ ⊥$}
    \end{bprooftree}
    \qquad
    \begin{bprooftree}
      \AxiomC{$Γ ⟝ φ ∧ ⊤$}
      \UnaryInfC{$Γ ⟝ φ$}
    \end{bprooftree}
    \qquad
    \begin{bprooftree}
      \AxiomC{$Γ ⟝ φ ∧ φ$}
      \UnaryInfC{$Γ ⟝ φ$}
    \end{bprooftree}
    \qquad
    \begin{bprooftree}
      \AxiomC{$Γ ⟝ φ ∧ ¬ φ$}
      \UnaryInfC{$Γ ⟝ ⊥$}
    \end{bprooftree}
  \]
  \caption{Theorems to remove inside of a formula by the
  \canonicalize rule.}
\label{fig:redundancies}
\end{figure}

The \canonicalize rule is mainly used by \Metis to introduce the negation of a
subgoal in its refutation proof. When the input formula is an axiom, definition
or hypothesis introduced in the \TPTP file of the problem, the \canonicalize
rule returns its negative normal form. In the remaining cases, as far as we
know, \Metis uses the \canonicalize rule to performs a sort of simplifications
to remove the tautologies listed in~Fig.~\ref{fig:redundancies} in the
conjunctive normal form of the input formula.

To reconstruct this rule, we guide its prensentation as follows. In
Lemma~\ref{lem:simplify-or} and Lemma~\ref{lem:simplify-and}, we show how to
simplify in disjunctions and conjunctions, respectively. The negative normal form
is stated in~Lemma~\ref{lem:nnf} and the conjunctive normal form in
Lemma~\ref{lem:cnf}. Finally, the reconstruction of the \canonicalize rule is
stated in Theorem~\ref{thm:canonicalize}.

% ----------------------------------------------------------------------------
% -- Simplify-Or --
% ----------------------------------------------------------------------------

\begin{notation}
In a disjunction, $φ ≡ φ₁ ∨ φ₂ ∨ \cdots ∨ φₙ$, we say $ψ ∈_{∨} φ$,
if there is some $i = 1, \cdots, n$ such that $ψ ≡ φᵢ$.
Note that $ψ ∈_{∨} φ$ is another representation for the equality
$ψ ≡ \freorder_{∨}~φ~ψ$.
We assume all disjunctions to be right-associative unless otherwise
stated.
\end{notation}

\begin{mainlemma}
  \label{lem:simplify-or}
  %Let $φ : \Prop$ be a right-associative formula.
  If $Γ~⟝~φ$ then $Γ~⟝~\fcanon_{∨}~φ$, where

\begin{equation}
\label{eq:canon-or}
\begin{aligned}
 &\hspace{.495mm}\fcanon_{∨} : \NProp \to \NProp\\
 &\begin{array}{lll}
   \fcanon_{∨} &(⊥ ∨ φ)     &= \fcanon_{∨}~φ \\
   \fcanon_{∨} &(φ ∨ ⊥)     &= \fcanon_{∨}~φ \\
   \fcanon_{∨} &(⊤ ∨ φ)     &= ⊤\\
   \fcanon_{∨} &(φ ∨ ⊤)     &= ⊤\\
   \fcanon_{∨} &(φ₁ ∨ φ₂)   &=
    \begin{cases}
     ⊤,                     & \text{ if } φ₁ ≡ ¬ ψ \text{ for some }ψ : \Prop \text{ and } ψ ∈_{∨} φ₂;\\
     ⊤,                     & \text{ if } (¬ φ₁) ∈_{∨} φ₂;\\
     \fcanon_{∨}~φ₂,        & \text{ if } φ₁ ∈_{∨} φ₂;\\
     ⊤,                     & \text{ if } \fcanon_{∨}~φ₂~ ≡ ⊤;\\
     φ₁,                    & \text{ if } \fcanon_{∨}~φ₂~ ≡ ⊥;\\
     φ₁ ∨ \fcanon_{∨}~φ₂,   & \text{ otherwise.}
    \end{cases}\\
   \fcanon_{∨} &φ           &= φ.
  \end{array}
\end{aligned}
\end{equation}
\end{mainlemma}

% ----------------------------------------------------------------------------
% -- Simplify-and --
% ----------------------------------------------------------------------------

Now, we have removed tautologies in disjunctions by applying the $\fcanon_{∨}$
function, we formulate in a similar way, the $\fcanon_{∧}$ function to simplify
tautologies about conjunctions.

\begin{notation}
In a conjunction, $φ ≡ φ₁ ∧ φ₂ ∧ \cdots ∧ φₙ$, we say
$ψ ∈_{∧} φ$, if there is some $i = 1, \cdots, n$ such that $ψ ≡ φᵢ$.
Note that $ψ ∈_{∧} φ$ is another representation of
the equality $ψ ≡ \fconjunct~φ~ψ$.
We assume all conjunctions to be right-associative unless otherwise
stated.
\end{notation}

\begin{mainlemma}
  \label{lem:simplify-and}
  %Let $φ : \Prop$ be a right-associative formula.
  If $Γ~⟝~φ$ then $Γ~⟝~\fcanon_{∧}~φ$, where

  \begin{equation}
   \label{eq:simplify-and}
    \begin{aligned}
     &\hspace{.495mm}\fcanon_{∧} : \NProp \to \NProp\\
      &\begin{array}{lll}
        \fcanon_{∧} &(⊥ ∧ φ)     &= ⊥  \\
        \fcanon_{∧} &(φ ∧ ⊥)     &= ⊥  \\
        \fcanon_{∧} &(⊤ ∧ φ)     &= \fcanon_{∧}~φ \\
        \fcanon_{∧} &(φ ∧ ⊤)     &= \fcanon_{∧}~φ \\
        \fcanon_{∧} &(φ₁ ∧ φ₂) &=
          \begin{cases}
            ⊥,                   & \text{ if } φ₁ ≡ ¬ ψ \text{ for some }ψ : \Prop \text{ and } ψ ∈_{∧} φ₂;\\
            ⊥,                   & \text{ if } (¬ φ₁) ∈_{∧} φ₂;\\
            \fcanon_{∧}~φ₂,      & \text{ if } φ₁ ∈_{∧} φ₂;\\
            φ₁,                  & \text{ if } \fcanon_{∧}~φ₂~≡ ⊤;\\
            ⊥,                   & \text{ if } \fcanon_{∧}~φ₂~≡ ⊥;\\
            φ₁ ∧ \fcanon_{∧}~φ₂, &\text{ otherwise.}
          \end{cases}\\
        \fcanon_{∧} &φ         &= φ.
       \end{array}
    \end{aligned}
    \end{equation}
\end{mainlemma}

Using the functions $\fcanon_{∨}$ and $\fcanon_{∧}$ we can define our function for
the negative normal form present in \Metis, we once again formulate this version
based on the \Metis source code to \emph{normalize} formulas. We use bounded
recursion to define the function $\fnnf_{1}$ in~\eqref{eq:nnf}.

% ----------------------------------------------------------------------------
% -- NNF --
% ----------------------------------------------------------------------------

@@

The function $\fnnf$ in~\eqref{eq:fnnf} is defined similarly as the \fstrip
function in~\eqref{eq:strip}, the complexity measure function $\fnnf_{cm}$ is
defined in Appendix in~\cite{Prieto-Cubides2017a}.

\begin{align}
  \label{eq:fnnf}
 \begin{split}
   &\fnnf : \NProp \to \NProp\\
   &\fnnf~φ = \fnnf_{1}~φ~(\fnnf_{cm} \varphi).
 \end{split}
\end{align}

\begin{mainlemma}
  \label{lem:nnf}
  If $Γ ⊢ φ$ then $Γ ⟝ \fnnf~φ$.
\end{mainlemma}

%%%
%%% TODO: 2018-06-25 17:14
%%%

@@

@@

### 2018-06-25 pubs/itp-2018/sections/proof-reconstruction/resolve.tex
Attempt to realocate the resolve rule after simplify

% -- TODO algun texto aqui?

\begin{mydefinition}

A \emph{literal} is a propositional variable (positive literal) or a
negation of a propositional variable (negative literal).

\end{mydefinition}

\begin{notation}
\Lit is a synonym of \Prop to refer to literals.
\end{notation}

\begin{mydefinition}

The \emph{negative normal form} of a formula is the logical
equivalent version of it in which negations appear only in the
literals and the formula does not contain any implications.

\end{mydefinition}

\begin{mydefinition}

The \emph{conjunctive normal form} of a formula also called clausal
normal form is the logical equivalent version expressed as a
conjunction of clauses where a \emph{clause} is the disjunction of
zero or more literals.
\end{mydefinition}

@@

The \resolve inference rule in combination with the \simplify rule is  the
resolution theorem for propositional logic that we found in the \Metis \TSTP
derivations (see this rule in~Fig.~\ref{fig:metis-inferences}). In particular,
to reconstruct the \resolve rule following the pattern in
Example~\ref{ex:inference-rule-pattern}, it is necessary to have a mechanism to
test for logic equivalence between propositions. The challenge is because of we
left out the semantics to treat only the syntax aspects of the propositional
formulas, therefore we have to manipulate in somehow the formulas to see if they
are logically equivalent.

Our approach to address this issue mainly consists of checking for syntactical
equality between the formulas. Firstly by converting them to their conjunctive
normal and secondly, by performing reordering of the inner formulas of the
resulting normal forms in the first conversion when they were not equal.

In~\cite{Prieto-Cubides2017a}, we make a description of the complete machinery
we use for that purpose. For instance, we define functions of the type ($\Source
\to  \Target \to \Prop$) like $\freorder_{∨}$ to fix the order of a disjunction
(premise) given by the order in the literals of another disjunction (conclusion)
if it is possible. Following the same way, we define for conjunctions the
function $\freorder_{\wedge}$ and for conjunctive normal forms the function
$\freorder_{\wedge\vee}$.

Using reordering after applying a customized version of the
resolution theorem defined in~\eqref{eq:rsol} we get the expected
result.

\begin{mainlemma}
  \label{lem:rsol}
  If $Γ ⊢ φ$ then $Γ ⊢ \frsol~φ$, where
  \begin{equation}
    \label{eq:rsol}
    \begin{aligned}
    &\frsol : \Prop \to \Prop\\
    &\begin{array}{ll}
      \frsol~φ &=
        \begin{cases}
          φ₂,      &\text{ if }φ ≡ (φ₁ ∨ φ₂) ∧ (¬ φ₁ ∨ φ₂);\\
          φ₂ ∨ φ₄, &\text{ if }φ ≡ (φ₁ ∨ φ₂) ∧ (¬ φ₁ ∨ φ₄);\\
          φ, &\text{ otherwise.}
        \end{cases}
      \end{array}
      \end{aligned}
\end{equation}
\end{mainlemma}

\begin{mainth}
  \label{thm:resolve}
  Let $ℓ$ be a literal, $ℓ : \Lit$, and $ψ : \Target$. If $Γ ⊢ φ₁$ and
  $Γ~⊢~φ₂$ then $Γ~⊢~\fresolve~φ₁~φ₂~ℓ~ψ$, where
  \begin{equation}
  \begin{split}
  \label{eq:resolve}
    &\fresolve : \Source \to \Source \to \Lit \to \Target \to \Prop\\
    &\fresolve~φ₁~φ₂~ℓ~ψ =
      \frsol~(\freorder_{∨}~φ₁~(ℓ ∨ ψ) ∧ \freorder_{∨}~φ₂~(¬ ℓ ∨ ψ)).
  \end{split}
  \end{equation}
\end{mainth}

>>>>

@@

The \resolve inference rule in combination with the \simplify rule are the
responsable rules to perform resolution. The \fresolve rules is a particular
version of the resolution theorem for propositional logic \TSTP derivations (see
this rule in~Fig.~\ref{fig:metis-inferences}).

\begin{example} In the \TSTP derivation in
Fig.~\ref{eq:complete-example-problem}, there are two ocurrences in
line~\ref{line14} and line~\ref{line16}~of the \resolve rule. This rule has two
inputs: the \emph{resolvent}, a literal denoted by ℓ in~Fig.\ref{fig:metis-inferences},
and a list of two formulas. \end{example}

In~\cite{Prieto-Cubides2017a}, we describe our first attempt to reconstruct this
rule that performs reordering of formulas jointly with the application of a
customized version of the resolution theorem to reconstruct the rule. We propose
an alternative definition formulated as a particular case
of the \fsimplify rule. We do this after seeing the behavior
of this rule in the \TSTP derivations from a more practical point of view for
the type-checking.

\begin{mainth}
  \label{thm:resolve}
  Let $ℓ$ be a literal, $ℓ : \Lit$, and $ψ : \Target$. If $Γ ⊢ φ₁$ and
  $Γ~⊢~φ₂$ then $Γ~⊢~\fresolve~φ₁~φ₂~ℓ~ψ$, where
  \begin{equation}
  \begin{split}
  \label{eq:resolve}
    &\fresolve : \Source \to \Source \to \Lit \to \Target \to \Prop\\
    &\fresolve~φ₁~φ₂~ℓ~ψ~=~\fsimplify~(φ₁ ∧ φ₂)~(¬ ℓ ∨ ℓ)~ψ
  \end{split}
  \end{equation}
\end{mainth}

%% maybe we want to show this easy proof.
% \begin{proof}
%   \begin{equation*}
%   \begin{bprooftree}
%     \AxiomC{$Γ ⊢ φ₁$}
%     \AxiomC{$Γ ⊢ φ₂$}
%     \RightLabel{∧-intro}
%     \BinaryInfC{$Γ ⊢~\varphi_1~\wedge~\varphi_2$}
%     \AxiomC{\abbre{PEM}~$\ell$}
%     \RightLabel{Theorem~\ref{thm:simplify}.}
%     \BinaryInfC{$\Gamma\vdash~\fsimplify~~(φ₁ ∧ φ₂)~(¬ ℓ ∨ ℓ)~ψ$}
%   \end{bprooftree}
%   \end{equation*}
% \end{proof}
